\bf Manifold Clustering in the Generalized Random Dot Product Graph

### 2022-04-22 draft/draft.Rmd
added proof of theorem 1

---
title: Manifold Clustering for Latent Structure Block Models

@@

\begin{definition}[Manifold block model]
Let $p, q \geq 0$, $d = p + q \geq 1$, $K \geq 1$, and $n \geq 1$ be integers.
If there are manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1] \}$, and 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}$ such that each $x_i \in \mathcal{M}_k$ for some $k \in [K]$, then $A \sim \GRDPG_{p,q}(X; \rho_n)$ is a \emph{manifold block model}.
\end{definition}

In practice, we define each manifold $\mathcal{M}_k$ by a function $f_k : [0, 1]^r \to \mathcal{X}$, where $1 \geq r < d$ is the dimensionality of manifold $\mathcal{M}_k$.

@@

\begin{example}
Let $p = 2$, $q = 0$, $K = 2$, and $r = 1$. 
Define two manifolds by $f_1(t) = \begin{bmatrix} \cos(\frac{\pi}{3} t) & \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$ and $f_2(t) = \begin{bmatrix} 1 - \cos(\frac{\pi}{3} t) & 1 - \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$.
Draw $t_1, ..., t_n \sim \Uniform(0, 1)$ and $z_1, ..., z_n \Multinomial(\frac{1}{2}, \frac{1}{2})$, and compute latent vectors $x_i = f_{z_i}(t_i)$, which are collected in data matrix $\begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$. 
Finally, let $A \sim \RDPG(X)$.

@@

par(mfrow = c(1, 3))
# par(fig = c(0, .7, 0, 1))
plot(data.matrix, col = z * 2, asp = 1,
     xlab = NA, ylab = NA, xaxt = 'n', yaxt = 'n',
     xlim = c(-.1, 1.1), ylim = c(0, 1))
# ggplot() +
#   geom_point(aes(x = data.matrix[, 1], y = data.matrix[, 2], colour = factor(z))) +
#   labs(x = NULL, y = NULL) +
#   scale_colour_brewer(palette = 'Set1') +
#   theme(legend.position = 'none') +
#   coord_fixed()
# par(fig = c(.3, 1, 0, 1))
qgraph::qgraph(A, groups = factor(z), legend = FALSE)
plot(Xhat, col = z * 2, asp = 1, 
     xlab = NA, ylab = NA, xaxt = 'n', yaxt = 'n')
```
\end{example}

@@

To prove theorems \ref{nonintersect-no-noise}, \ref{nonintersect-with-noise}, and \ref{nonintersect-mbm}, we first need to introduce some theory for order statistics. 
Let $D_i = X_{(i+1)} - X_{(i)}$. Then if $\max_i D_i < \delta$, we have sufficient separation of points in $\mathcal{M}_1$. 
Then it is sufficient to quantify $P(\max_i D_i > \delta)$ as a function of $n$ and $\delta$ and show that this converges to zero as $n$ grows to $\infty$.

We denote $f(x)$ as the density of each $X_i$, $g_i(x)$ as the density of $X_{(i)}$, 
$g_{ij}(x, y)$ as the joint density of $X_{(i)}, X_{(j)}$, and $h_i(d)$ as the density of $D_i$ (with corresponding capital letters for the cumulative distribution functions).

@@

\begin{lemma}[The probability density function of $D_i$]
\begin{equation}
\label{eq:pdf}
h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \dd x 
\end{equation}
\end{lemma}

\begin{proof}
This is just a direct consequence of 2 and 3 under the given statements. 
We also note that because the support of $X_i$ is $[0, 1]$, 
the integral only needs to be evaluated from $0$ to $1 - d$ 
because of the $f(x+d)$ and $1 - F(x+d)$ terms.
\end{proof}

\begin{lemma}[The cumulative distribution function of $D_i$]

@@

Because of the $x + \delta$ term, we can't actually evaluate this integral all the way up to $1$, and so we are left with 
$$
\begin{aligned}
& = 1 - \int_0^{1 - \delta} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x.
\end{aligned}
$$
\end{proof}

@@

$$
h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \dd x
$$

@@

$$
= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \dd x
$$

@@

$$
\begin{aligned}
& = \frac{n!}{(i-1)! (n-i-1)!} \frac{(i-1)!}{(n-i) \cdots (n-2)} \int_0^{1-d} (1-x-d)^{n-2} \dd x \\
& = -\frac{n (n-1)}{n-1} (1-x-d)^{n-1} \Big|_0^{1-d} \\
& = n (1 - d)^{n-1}
\end{aligned}
$$

@@

\begin{proof}[Proof (sketch)]
Since $X_{(i+1)} - X_{(i)} = D_i \sim \Betadist(1, n)$, 
$P(X_{(i+1)} - X_{(i)} < \delta) = 1 - (1 - \delta)^n $. This yields

@@

\begin{lemma}
Let $X_1, ..., X_n \iid F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then for any $\epsilon > 0$, there exists $N > 0$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
\end{lemma}

\begin{proof}[Proof (sketch)]
We start with Eq. (\ref{eq:cdf}):

@@

Here, we extend the results on the line to the unit hypercube. The following theorem is a direct consequence of Lemma 2 of \citet{trosset2020rehabilitating}.

@@

>>>>

---
title: Manifold Clustering in the Generalized Random Dot Product Graph

@@

\begin{definition}[Manifold block model]
Let $p, q \geq 0$, $d = p + q \geq 1$, $K \geq 1$, and $n \geq 1$ be integers.
If there are manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1] \}$, and 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ such that each $x_i \in \mathcal{M}_k$ for some $k \in [K]$, then $A \sim \GRDPG_{p,q}(X; \rho_n)$ is a \emph{manifold block model}.
\end{definition} 
In practice, we define each manifold $\mathcal{M}_k$ by a continuous function $g_k : [0, 1]^r \to \mathcal{X}$, where $1 \leq r < d$ is the dimensionality of manifold $\mathcal{M}_k$.

@@

\begin{example}
Let $p = 2$, $q = 0$, $K = 2$, and $r = 1$. 
Define two manifolds by $f_1(t) = \begin{bmatrix} \cos(\frac{\pi}{3} t) & \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$ and $f_2(t) = \begin{bmatrix} 1 - \cos(\frac{\pi}{3} t) & 1 - \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$.
Draw $t_1, ..., t_n \sim \Uniform(0, 1)$ and $z_1, ..., z_n \iid \Multinomial(\frac{1}{2}, \frac{1}{2})$, and compute latent vectors $x_i = f_{z_i}(t_i)$, which are collected in data matrix $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$. 
Finally, let $A \sim \RDPG(X)$.

@@

par(mfrow = c(1, 3))
# par(fig = c(0, .7, 0, 1))
plot(data.matrix, col = z * 2, asp = 1,
     xlab = NA, ylab = NA, 
     # xaxt = 'n', yaxt = 'n',
     xlim = c(-.1, 1.1), ylim = c(0, 1))
# ggplot() +
#   geom_point(aes(x = data.matrix[, 1], y = data.matrix[, 2], colour = factor(z))) +
#   labs(x = NULL, y = NULL) +
#   scale_colour_brewer(palette = 'Set1') +
#   theme(legend.position = 'none') +
#   coord_fixed()
# par(fig = c(.3, 1, 0, 1))
qgraph::qgraph(A, groups = factor(z), legend = FALSE)
plot(Xhat, col = z * 2, asp = 1, 
     # xaxt = 'n', yaxt = 'n', 
     xlab = NA, ylab = NA)
```
\end{example}

@@

To prove theorems \ref{nonintersect-no-noise}, \ref{nonintersect-with-noise}, and \ref{nonintersect-mbm}, we first need to introduce some theory for order statistics. 
Let $D_i^{(k)} = X_{(i+1)}^{(k)} - X_{(i)}^{(k)}$, where $X_{(i)}^{(k)}$ is the $i^{th}$ order statistic of univariate sample $X_1^{(k)}, ..., X_n^{(k)} \iid F^{(k)}$ and $F^{(k)}$ is a continuous distribution on manifold $\mathcal{M}_k$. 
If $\max_i D_i^{(k)} < \delta$, then there is sufficient separation of points between each manifold. 
Then it is sufficient to quantify $P(\max_i D_i^{(k)} > \delta)$ for each $k$ as a function of $n$ and $\delta$ and show that this converges to zero as $n$ grows to $\infty$.

For ease of notation, we drop the superscript denoting community/manifold membership until lemma 6. 
Denote $f(x)$ as the density of each $F$, $g_i(x)$ as the density of $X_{(i)}$, 
$g_{ij}(x, y)$ as the joint density of $X_{(i)}, X_{(j)}$, and $h_i(d)$ as the density of $D_i$ (with corresponding capital letters for the cumulative distribution functions).

@@

\begin{lemma}[The probability density function of $D_i$]
Suppose $X_1, ..., X_n \iid F$ and the support of $F$ is the unit interval. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then the density of each $D_i$ is:

\begin{equation}
\label{eq:pdf}
h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \dd x 
\end{equation}
\end{lemma}

\begin{proof}
This is a direct consequence of 2 and 3. 
We also note that because the support of $X_i$ is $[0, 1]$, 
the integral only needs to be evaluated from $0$ to $1 - d$ 
because of the $f(x+d)$ and $1 - F(x+d)$ terms.
\end{proof}

\begin{lemma}[The cumulative distribution function of $D_i$]
Suppose $X_1, ..., X_n \iid F$ and the support of $F$ is the unit interval. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then the distribution function of each $D_i$ is:

@@

Because of the $x + \delta$ term and the fact that the support is $[0, 1]$, the integrand is zero above $1 - \delta$, so we are left with

$$
\begin{aligned}
& = 1 - \int_0^{1 - \delta} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x.
\end{aligned}
$$
\end{proof}

@@

$$
h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \dd x.
$$

@@

$$
= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \dd x.
$$

@@

$$
\begin{aligned}
& = \frac{n!}{(i-1)! (n-i-1)!} \frac{(i-1)!}{(n-i) \cdots (n-2)} \int_0^{1-d} (1-x-d)^{n-2} \dd x \\
& = -\frac{n (n-1)}{n-1} (1-x-d)^{n-1} \Big|_0^{1-d} \\
& = n (1 - d)^{n-1}.
\end{aligned}
$$

@@

\begin{proof}
Since $X_{(i+1)} - X_{(i)} = D_i \sim \Betadist(1, n)$, 
$P(X_{(i+1)} - X_{(i)} < \delta) = 1 - (1 - \delta)^n $. This yields

@@

\begin{lemma}
\label{lem:lim-d}
Let $X_1, ..., X_n \iid F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then for any $\epsilon > 0$, there exists $N > 0$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
\end{lemma}

\begin{proof}
We start with Eq. (\ref{eq:cdf}):

@@

\begin{proof}[Proof of Theorem \ref{nonintersect-no-noise}]
For this theorem to be true, there are two additional conditions: 
First, each $n_k \to \infty$ as $n \to \infty$, or the sample size on each manifold grows with $n$, and second, the number of manifolds $K$ remains fixed. 
Then this is a direct consequence of Lemma \ref{lem:lim-d}. 
Here, we denote $m = \min_k n_k$ as the sample size of the manifold with the fewest points.

$$
\begin{aligned}
P(\max_{k,i} \|X_{(i+1)}^{(k)} - X_{(i)}^{(k)}\| < \delta) & = P(\forall k, i ~ D_i^{(k)} < \delta) \\
& \geq \prod_{k=1}^K P(D_i^{(k)} < \delta) \\
& \geq (1 - (m-1) (1 - a \delta)^{m} + (m-1) (1 - F(1-\delta) - a \delta)^{m})^K
\end{aligned}
$$
Since as $n \to \infty$, $m \to \infty$ as well, and $K$ is fixed, this approaches $(1)^K = 1$. 
\end{proof}

Here, we extend the results on the line to the unit hypercube. The following theorem is based on Lemma 2 of \citet{trosset2020rehabilitating}.

@@

\section{Details on Fitting Bezier Curves with Noise}

### 2022-04-22 draft/draft.tex
added proof of theorem 1

\if0\blind
{
  \title{\bf Manifold Clustering for Latent Structure Block Models}

@@

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Manifold Clustering for Latent Structure Block Models}
  \end{center}
  \medskip
} \fi

@@

\begin{definition}[Manifold block model]
Let $p, q \geq 0$, $d = p + q \geq 1$, $K \geq 1$, and $n \geq 1$ be integers.
If there are manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1] \}$, and 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}$ such that each $x_i \in \mathcal{M}_k$ for some $k \in [K]$, then $A \sim \mathrm{GRDPG}_{p,q}(X; \rho_n)$ is a \emph{manifold block model}.
\end{definition}

In practice, we define each manifold \(\mathcal{M}_k\) by a function
\(f_k : [0, 1]^r \to \mathcal{X}\), where \(1 \geq r < d\) is the
dimensionality of manifold \(\mathcal{M}_k\).

@@

\begin{example}
Let $p = 2$, $q = 0$, $K = 2$, and $r = 1$. 
Define two manifolds by $f_1(t) = \begin{bmatrix} \cos(\frac{\pi}{3} t) & \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$ and $f_2(t) = \begin{bmatrix} 1 - \cos(\frac{\pi}{3} t) & 1 - \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$.
Draw $t_1, ..., t_n \sim \mathrm{Uniform}(0, 1)$ and $z_1, ..., z_n \mathrm{Multinomial}(\frac{1}{2}, \frac{1}{2})$, and compute latent vectors $x_i = f_{z_i}(t_i)$, which are collected in data matrix $\begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$. 
Finally, let $A \sim \mathrm{RDPG}(X)$.

@@

To prove theorems \ref{nonintersect-no-noise},
\ref{nonintersect-with-noise}, and \ref{nonintersect-mbm}, we first need
to introduce some theory for order statistics. Let
\(D_i = X_{(i+1)} - X_{(i)}\). Then if \(\max_i D_i < \delta\), we have
sufficient separation of points in \(\mathcal{M}_1\). Then it is
sufficient to quantify \(P(\max_i D_i > \delta)\) as a function of \(n\)
and \(\delta\) and show that this converges to zero as \(n\) grows to
\(\infty\).

We denote \(f(x)\) as the density of each \(X_i\), \(g_i(x)\) as the
density of \(X_{(i)}\), \(g_{ij}(x, y)\) as the joint density of
\(X_{(i)}, X_{(j)}\), and \(h_i(d)\) as the density of \(D_i\) (with
corresponding capital letters for the cumulative distribution
functions).

@@

\begin{lemma}[The probability density function of $D_i$]
\begin{equation}
\label{eq:pdf}
h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \mathrm{d}x 
\end{equation}
\end{lemma}

\begin{proof}
This is just a direct consequence of 2 and 3 under the given statements. 
We also note that because the support of $X_i$ is $[0, 1]$, 
the integral only needs to be evaluated from $0$ to $1 - d$ 
because of the $f(x+d)$ and $1 - F(x+d)$ terms.
\end{proof}

\begin{lemma}[The cumulative distribution function of $D_i$]

@@

Because of the $x + \delta$ term, we can't actually evaluate this integral all the way up to $1$, and so we are left with 
$$
\begin{aligned}
& = 1 - \int_0^{1 - \delta} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x.
\end{aligned}
$$
\end{proof}

@@

$$
h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \mathrm{d}x
$$

@@

$$
= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \mathrm{d}x
$$

@@

$$
\begin{aligned}
& = \frac{n!}{(i-1)! (n-i-1)!} \frac{(i-1)!}{(n-i) \cdots (n-2)} \int_0^{1-d} (1-x-d)^{n-2} \mathrm{d}x \\
& = -\frac{n (n-1)}{n-1} (1-x-d)^{n-1} \Big|_0^{1-d} \\
& = n (1 - d)^{n-1}
\end{aligned}
$$

@@

\begin{proof}[Proof (sketch)]
Since $X_{(i+1)} - X_{(i)} = D_i \sim \mathrm{Beta}(1, n)$, 
$P(X_{(i+1)} - X_{(i)} < \delta) = 1 - (1 - \delta)^n $. This yields

@@

\begin{lemma}
Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then for any $\epsilon > 0$, there exists $N > 0$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
\end{lemma}

\begin{proof}[Proof (sketch)]
We start with Eq. (\ref{eq:cdf}):

>>>>

\if0\blind
{
  \title{\bf Manifold Clustering in the Generalized Random Dot Product
Graph}

@@

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Manifold Clustering in the Generalized Random Dot Product
Graph}
  \end{center}
  \medskip
} \fi

@@

\begin{definition}[Manifold block model]
Let $p, q \geq 0$, $d = p + q \geq 1$, $K \geq 1$, and $n \geq 1$ be integers.
If there are manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1] \}$, and 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ such that each $x_i \in \mathcal{M}_k$ for some $k \in [K]$, then $A \sim \mathrm{GRDPG}_{p,q}(X; \rho_n)$ is a \emph{manifold block model}.
\end{definition}

In practice, we define each manifold \(\mathcal{M}_k\) by a continuous
function \(g_k : [0, 1]^r \to \mathcal{X}\), where \(1 \leq r < d\) is
the dimensionality of manifold \(\mathcal{M}_k\).

@@

\begin{example}
Let $p = 2$, $q = 0$, $K = 2$, and $r = 1$. 
Define two manifolds by $f_1(t) = \begin{bmatrix} \cos(\frac{\pi}{3} t) & \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$ and $f_2(t) = \begin{bmatrix} 1 - \cos(\frac{\pi}{3} t) & 1 - \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$.
Draw $t_1, ..., t_n \sim \mathrm{Uniform}(0, 1)$ and $z_1, ..., z_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\frac{1}{2}, \frac{1}{2})$, and compute latent vectors $x_i = f_{z_i}(t_i)$, which are collected in data matrix $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$. 
Finally, let $A \sim \mathrm{RDPG}(X)$.

@@

To prove theorems \ref{nonintersect-no-noise},
\ref{nonintersect-with-noise}, and \ref{nonintersect-mbm}, we first need
to introduce some theory for order statistics. Let
\(D_i^{(k)} = X_{(i+1)}^{(k)} - X_{(i)}^{(k)}\), where \(X_{(i)}^{(k)}\)
is the \(i^{th}\) order statistic of univariate sample
\(X_1^{(k)}, ..., X_n^{(k)} \stackrel{\mathrm{iid}}{\sim}F^{(k)}\) and
\(F^{(k)}\) is a continuous distribution on manifold \(\mathcal{M}_k\).
If \(\max_i D_i^{(k)} < \delta\), then there is sufficient separation of
points between each manifold. Then it is sufficient to quantify
\(P(\max_i D_i^{(k)} > \delta)\) for each \(k\) as a function of \(n\)
and \(\delta\) and show that this converges to zero as \(n\) grows to
\(\infty\).

For ease of notation, we drop the superscript denoting
community/manifold membership until lemma 6. Denote \(f(x)\) as the
density of each \(F\), \(g_i(x)\) as the density of \(X_{(i)}\),
\(g_{ij}(x, y)\) as the joint density of \(X_{(i)}, X_{(j)}\), and
\(h_i(d)\) as the density of \(D_i\) (with corresponding capital letters
for the cumulative distribution functions).

@@

\begin{lemma}[The probability density function of $D_i$]
Suppose $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ and the support of $F$ is the unit interval. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then the density of each $D_i$ is:

\begin{equation}
\label{eq:pdf}
h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \mathrm{d}x 
\end{equation}
\end{lemma}

\begin{proof}
This is a direct consequence of 2 and 3. 
We also note that because the support of $X_i$ is $[0, 1]$, 
the integral only needs to be evaluated from $0$ to $1 - d$ 
because of the $f(x+d)$ and $1 - F(x+d)$ terms.
\end{proof}

\begin{lemma}[The cumulative distribution function of $D_i$]
Suppose $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ and the support of $F$ is the unit interval. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then the distribution function of each $D_i$ is:

@@

Because of the $x + \delta$ term and the fact that the support is $[0, 1]$, the integrand is zero above $1 - \delta$, so we are left with

$$
\begin{aligned}
& = 1 - \int_0^{1 - \delta} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x.
\end{aligned}
$$
\end{proof}

@@

$$
h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \mathrm{d}x.
$$

@@

$$
= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \mathrm{d}x.
$$

@@

$$
\begin{aligned}
& = \frac{n!}{(i-1)! (n-i-1)!} \frac{(i-1)!}{(n-i) \cdots (n-2)} \int_0^{1-d} (1-x-d)^{n-2} \mathrm{d}x \\
& = -\frac{n (n-1)}{n-1} (1-x-d)^{n-1} \Big|_0^{1-d} \\
& = n (1 - d)^{n-1}.
\end{aligned}
$$

@@

\begin{proof}
Since $X_{(i+1)} - X_{(i)} = D_i \sim \mathrm{Beta}(1, n)$, 
$P(X_{(i+1)} - X_{(i)} < \delta) = 1 - (1 - \delta)^n $. This yields

@@

\begin{lemma}
\label{lem:lim-d}
Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then for any $\epsilon > 0$, there exists $N > 0$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
\end{lemma}

\begin{proof}
We start with Eq. (\ref{eq:cdf}):
